### 2018-11-15 Manuscript/Manuscript.tex

To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (Macdonald et al., in review; Fig. \ref{fig:LIP_area}E). The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state. The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals. Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states. We take two approaches for comparison between the LIP area reconstruction and the record of ice extent. The first is to analyze the correlation between LIP area and the extent of ice away from the pole (Fig. \ref{fig:LIP_correlation}). The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles). This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.

@@

\begin{table}[h!]
\scriptsize
\caption{Phanerozoic large igneous provinces (and the Franklin).}
\resizebox{\textwidth}{!}{
\begin{tabular}{lc>{\raggedright}p{4cm}cc>{\raggedright}p{4cm}ccl}
\toprule
\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original} & \textbf{present} & \textbf{present area ref.} & \textbf{present/} & \textbf{half-life}$^{1}$ & \textbf{buried?} \\
 & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
 & & & \textbf{(Mkm$^{2}$)} & \textbf{(Mkm$^{2}$)} & & & & \\
\midrule
Columbia River & 16 & \cite{Kasbohm2018a} & 0.68 & 0.38 & \cite{Buchan2004a} & 0.56 & 19.2 & no \\
Afar & 30 & \cite{Courtillot2003a} & 2.05 & 0.63 & \cite{Coffin2006a} & 0.31 & 17.7 & partial \\
NAIP & 62 & \cite{Larsen2015a} & 1.07 & 0.29 & \cite{Buchan2004a, Coffin2006a} & 0.27 & 33.0 & partial \\
Deccan & 66 & \cite{Schoene2014a} & 0.83 & 0.56 & \cite{Coffin2006a} & 0.68 & 116.6 & no \\
Seychelles & 66 & \cite{Schoene2014a} & 0.46 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
Madagascar & 90 & \cite{Cucciniello2010a} & 0.63 & 0.03 & \cite{Coffin2006a} & 0.05 & 20.8 & no \\
Caribbean-Colombian & 94 & \cite{Loewen2013a} & 0.71 & 0.13 & \cite{Coffin2006a} & 0.18 & 37.6 & no \\
HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Moosdorf2015a} & 0.04 & 20.7 & no \\
EQUAMP & 131 & \cite{Hollanda2016a} & 0.66 & 0.01 & \cite{Hollanda2016a} & 0.01 & 20.5 & no \\
Comei & 132 & \cite{Zhu2009a} & 0.11 & - & - & - & - & no \\
Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\
NW Australia Margin & 160 & \cite{Pirajno2012a} & 0.62 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{2}$ & 0.05 & 41.3 & no \\
Ferrar & 183 & \cite{Burgess2015a} & 0.18 & - & - & - & - & no \\
CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{3}$ & 0.02 & 35.7 & partial \\
Siberia & 252 & \cite{Burgess2015b} & 3.46 & 0.47 & \cite{Coffin2006a} & 0.14 & 87.5 & no \\
Emeishan & 259 & \cite{Zhou2002a} & 0.71 & 0.06 & \cite{Coffin2006a} & 0.09 & 72.9 & no \\
Panjal-Qiangtang & 283 & \cite{Zhai2013a} & 0.11 & - & - & - & - & no \\
Tarim & 290 & \cite{Xu2014a} & 0.35 & - & - & - & - & no \\
Magdalen & 360 & \cite{Brendan_Murphy1999a} & 0.42 & - & - & - & - & no \\
Vilyui & 374 & \cite{Ricci2013a} & 1.14 & - & - & - & - & no \\
Kola-Dnieper & 380 & \cite{Arzamastsev2014a} & 5.90 & - & - & - & - & no \\
Suordakh & 450 & \cite{Khudoley2013a} & 0.02 & - & - & - & - & no \\
Kalkarindji & 511 & \cite{Jourdan2014a} & 3.54 & 0.17 & \cite{Thorne2014a} & 0.05 & 116.3 & no \\
Franklin & 720 & \cite{Denyszyn2009a} & 2.62 & 0.04 & \cite{Buchan2004a} & 0.02 & 121.8 & no \\
\bottomrule
\end{tabular}
}
\begin{tablenotes}[para,flushleft]
$^{1}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.

>>>>

To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2018a}; Fig. \ref{fig:LIP_area}E). The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state. The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals. Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states. We take two approaches for comparison between the LIP area reconstruction and the record of ice extent. The first is to analyze the correlation between LIP area and the extent of ice away from the pole (Fig. \ref{fig:LIP_correlation}). The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles). This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.

@@

\begin{table}[h!]
\scriptsize
\caption{Phanerozoic large igneous provinces (and the Franklin).}
\resizebox{\textwidth}{!}{
\begin{tabular}{lc>{\raggedright}p{4cm}cc>{\raggedright}p{4cm}ccl}
\toprule
\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original} & \textbf{present} & \textbf{present area ref.} & \textbf{present/} & \textbf{half-life}$^{1}$ & \textbf{buried?} \\
 & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
 & & & \textbf{(Mkm$^{2}$)} & \textbf{(Mkm$^{2}$)} & & & & \\
\midrule
Columbia River & 16 & \cite{Kasbohm2018a} & 0.68 & 0.38 & \cite{Buchan2004a} & 0.56 & 19.2 & no \\
Afar & 30 & \cite{Courtillot2003a} & 2.05 & 0.63 & \cite{Coffin2006a} & 0.31 & 17.7 & partial \\
NAIP & 62 & \cite{Larsen2015a} & 1.07 & 0.29 & \cite{Buchan2004a, Coffin2006a} & 0.27 & 33.0 & partial \\
Deccan & 66 & \cite{Schoene2014a} & 0.83 & 0.56 & \cite{Coffin2006a} & 0.68 & 116.6 & no \\
Seychelles & 66 & \cite{Schoene2014a} & 0.46 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
Madagascar & 90 & \cite{Cucciniello2010a} & 0.63 & 0.03 & \cite{Coffin2006a} & 0.05 & 20.8 & no \\
Caribbean-Colombian & 94 & \cite{Loewen2013a} & 0.71 & 0.13 & \cite{Coffin2006a} & 0.18 & 37.6 & no \\
HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Moosdorf2015a} & 0.04 & 20.7 & no \\
EQUAMP & 131 & \cite{Hollanda2016a} & 0.66 & 0.01 & \cite{Hollanda2016a} & 0.01 & 20.5 & no \\
Comei & 132 & \cite{Zhu2009a} & 0.11 & - & - & - & - & no \\
Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018a} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\
NW Australia Margin & 160 & \cite{Pirajno2012a} & 0.62 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{2}$ & 0.05 & 41.3 & no \\
Ferrar & 183 & \cite{Burgess2015a} & 0.18 & - & - & - & - & no \\
CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{3}$ & 0.02 & 35.7 & partial \\
Siberia & 252 & \cite{Burgess2015b} & 3.46 & 0.47 & \cite{Coffin2006a} & 0.14 & 87.5 & no \\
Emeishan & 259 & \cite{Zhou2002a} & 0.71 & 0.06 & \cite{Coffin2006a} & 0.09 & 72.9 & no \\
Panjal-Qiangtang & 283 & \cite{Zhai2013a} & 0.11 & - & - & - & - & no \\
Tarim & 290 & \cite{Xu2014a} & 0.35 & - & - & - & - & no \\
Magdalen & 360 & \cite{Brendan_Murphy1999a} & 0.42 & - & - & - & - & no \\
Vilyui & 374 & \cite{Ricci2013a} & 1.14 & - & - & - & - & no \\
Kola-Dnieper & 380 & \cite{Arzamastsev2014a} & 5.90 & - & - & - & - & no \\
Suordakh & 450 & \cite{Khudoley2013a} & 0.02 & - & - & - & - & no \\
Kalkarindji & 511 & \cite{Jourdan2014a} & 3.54 & 0.17 & \cite{Thorne2014a} & 0.05 & 116.3 & no \\
Franklin & 720 & \cite{Denyszyn2009a} & 2.62 & 0.04 & \cite{Buchan2004a} & 0.02 & 121.8 & no \\
\bottomrule
\end{tabular}
}
\begin{tablenotes}[para,flushleft]
$^{1}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.
